Enforced Invariant Confluence

### 2018-04-21 doc/paper/sections/evaluation.tex
Added case studies to eval.

\section{Evaluation}\seclabel{Evaluation}
In this section, we describe Lucy: an implementation of our decision procedures
and system models. We then evaluate Lucy by answering the following questions:
\begin{itemize}
  \item
    How practical is the interactive invariant-confluence decision procedure?
    Can we use it to classify real-world transactions and invariants?
  \item
    How practical is segmented invariant-confluence? Are real-world workloads
    amenable to segmentation?
  \item
    How efficient is the interactive invariant-confluence decision procedure?
  \item
    How efficiently can we replicate a segmented invariant-confluent object as
    compared to alternative approaches like replicating with weak or strong
    consistency?
  \item
    How does the performance of replicating a segmented invariant-confluence
    object vary as we vary the workload, segmentation, and replication factor?
\end{itemize}

\subsection{Implementation}
Lucy includes an implementation of the interactive decision procedure described
in \algoref{InteractiveDecisionProcedure}, an implementation of a decision
procedure which checks criteria (1) - (4) from \thmref{LatticeProperty}, and an
implementation of the decision procedure described in
\algoref{ArbitraryStartInteractiveDecisionProcedure}. The decision procedures
are implemented in roughly 2,500 lines of Python. Users specify objects,
transactions, and invariants in a small Python DSL and interact with the
interactive decision procedures using an interactive Python console. We use
Z3~\cite{de2008z3} to implement our invariant-closure decision procedure,
compiling objects and invariants into a formula which is satisfiable if and
only if the object is \emph{not} invariant-closed. If the object is
invariant-closed, then Z3 concludes that the formula is unsatisfiable.
Otherwise, if the object is not invariant-closed, then Z3 produces a
counterexample witnessing the satisfiability of the formula.

@@

\subsection{Decision Procedures}
% \todo{Bank account with and without decrements.}
% \todo{Auction example.}
% \todo{Foreign keys with and without deletions.}
% \subsection{TPC and Feral}
% \todo{See which of the Feral and TPC transactions we can determine automatically}

\subsection{Segmented Invariant Confluence}%
\seclabel{SegmentedInvariantConfluenceEval}
In this section, we evaluate the performance of replicating an object with
segmented invariant-confluence as compared to replicating it with eventual
consistency or linearizability. We begin with two benchmarks that demonstrate
the same concept: the performance of segmented invariant-confluent replication
varies with the amount of global coordination induced by either (a) performing
a transaction that is disallowed within a segment or (b) transitioning between
segments.

>>>>

\section{Evaluation}\seclabel{Evaluation}
In this section, we describe and evaluate Lucy: a prototype implementation of
our decision procedures and system models.
% We evaluate Lucy by answering the following questions:
% \begin{itemize}
%   \item
%     How practical is the interactive invariant-confluence decision procedure?
%     Can we use it to classify real-world transactions and invariants?
%   \item
%     How practical is segmented invariant-confluence? Are real-world workloads
%     amenable to segmentation?
%   \item
%     How efficient is the interactive invariant-confluence decision procedure?
%   \item
%     How efficiently can we replicate a segmented invariant-confluent object as
%     compared to alternative approaches like replicating with weak or strong
%     consistency?
%   \item
%     How does the performance of replicating a segmented invariant-confluence
%     object vary as we vary the workload, segmentation, and replication factor?
% \end{itemize}

\subsection{Implementation}
Lucy includes an implementation of the interactive decision procedure described
in \algoref{InteractiveDecisionProcedure}, an implementation of a decision
procedure which checks criteria (1) - (4) from \thmref{LatticeProperty}, and an
implementation of the decision procedure described in
\algoref{ArbitraryStartInteractiveDecisionProcedure}. The decision procedures
are implemented in roughly 2,500 lines of Python. Users specify objects,
transactions, and invariants in a small Python DSL and interact with the
interactive decision procedures using an interactive Python console. We use
Z3~\cite{de2008z3} to implement our invariant-closure decision procedure,
compiling an object and invariant into a formula that is satisfiable if and
only if the object is \emph{not} invariant-closed. If the object is
invariant-closed, then Z3 concludes that the formula is unsatisfiable.
Otherwise, if the object is not invariant-closed, then Z3 produces a
counterexample witnessing the satisfiability of the formula.

@@

\subsection{Decision Procedures}
We now evaluate the practicality and efficiency of our decision procedure
prototypes. We begin by demonstrating the decision procedure on a handful of
simple, yet practical examples. We then, discuss TODO

% \todo{Bank account with and without decrements.}
% \todo{Auction example.}
% \todo{Foreign keys with and without deletions.}
% \subsection{TPC and Feral}
% \todo{See which of the Feral and TPC transactions we can determine automatically}
\example[$\ints^2$]\examplelabel{TwoIntsEval}
We begin with a minimal working example. Consider again our recurring example
of $\ints^2$ from \exampleref{Z2}. The Python code used to describe the object,
transactions, and invariant is given in \figref{Z2Code}. When we call
\python{checker.check()}, the interactive decision procedure produces
a counterexample in less than a tenth of a second.  After we label the
counterexample and refine the invariant with $y \leq 0$, the interactive
decision procedure determines that the object is invariant-confluent, again, in
less than a tenth of a second. Note that the object is invariant-confluent but
\emph{not} invariant-closed, so prior work that relies on invariant-closure
alone to determine invariant-confluence would not be able to identify this
example as invariant-confluent.

\begin{figure}[ht]
  \begin{Python}[gobble=4]
    checker = InteractiveInvariantConfluenceChecker()
    x = checker.int_max('x', 0) # An int, x, merged by max.
    y = checker.int_max('y', 0) # An int, y, merged by max.
    checker.add_transaction('increment_x', [x.assign(x + 1)])
    checker.add_transaction('decrement_y', [y.assign(y - 1)])
    checker.add_invariant(x * y <= 0)
    checker.check()
  \end{Python}
  \caption{\exampleref{TwoIntsEval} specification}\figlabel{Z2Code}
\end{figure}

\example[Foreign Keys]\examplelabel{ForeignKeysEval}
A 2P-Set $X = (A_X, R_X)$ is a set CRDT composed of a set of additions $A_X$
and a set of removals $R_X$~\cite{shapiro2011comprehensive}. We view the state
of the set $X$ as the difference $A_X - R_X$ of the addition and removal sets.
To add an element $x$ to the set, we add $x$ to $A_X$. Similarly, to remove $x$
from the set, we add it to $R_X$. The merge of two 2P-sets is a pairwise union
(i.e. $(A_X, R_X) \join (A_Y, R_Y) = (A_X \cup A_Y, R_X \cup R_Y)$).

We can use 2P-sets to model a simple relational database with foreign key
constraints. Let object $O = (X, Y) = ((A_X, R_X), (A_Y, R_Y))$ consist of a
pair of two 2P-Sets $X$ and $Y$, which we view as relations. Our invariant $X
\subseteq Y$ (i.e. $(A_X - R_X) \subseteq (A_Y - R_Y)$) models a foreign key
constraint from $X$ to $Y$. We ran our decision procedure on the object with
initial state $((\emptyset, \emptyset), (\emptyset, \emptyset))$ and
transactions that allow arbitrary insertions and deletions into $X$ and $Y$.
After less than a tenth of a second, the decision procedure produced a
reachable counterexample witnessing the fact that the object is not
invariant-confluent. A concurrent insertion into $X$ and deletion from $Y$ can
lead to a state which violates the invariant. This object is not
invariant-confluent and therefore not invariant-closed. Thus, previous tools
depending on invariant-closure as a sufficient condition would be unable to
conclude definitively that the object is \emph{not} invariant-confluent.

We also reran the decision procedure, but this time with insertions into $X$
and deletions from $Y$ disallowed. In less than a tenth of a second, the
decision procedure correctly deduced that the object is now
invariant-confluent. These results were manually proven
in~\cite{bailis2014coordination}, but our tool was able to confirm them
automatically in a negligible amount of time.

\example[Auction]\examplelabel{AuctionEval}
We now consider a simple auction system introduced in~\cite{gotsman2016cause}.
Our object consists of a set $B$ of integer-valued bids and a optional winning
bid $w$. Initially, $B = \emptyset$ and $w = \bot$ (indicating that there is no
winning bid yet) and we merge states by taking the union of $B$ and the maximum
of $w$ (where $\bot < n$ for all integers $n$). One transaction $t_b$ places a bid
$b$ by inserting it into $B$. Another transaction $t_\text{close}$ closes the
auction and sets $w$ equal to the largest bid in $B$. Our invariant is that if
the auction is closed (i.e.\ $w \neq \bot$), then $w = \max(B)$. We ran our
decision procedure on this example and in a third of a second, it produced a
reachable counterexample witnessing the fact that the object is not
invariant-confluent.  If we concurrently close the auction and place an large
bid, then we can end up in a state in which the auction is closed, but there is
a higher bid in $B$.

We then segmented our object as follows. The first segment $(\setst{(B, w)}{w =
\bot}, \setst{t_b}{b \in \ints})$ allows any bidding transaction so long as the
bid is open. The second segment $(\setst{B, w}{I \cap w \neq \bot}, \emptyset)$
includes all auctions which have already been closed and forbids all
transactions. This segmentation captures the intuition that bids should be
permitted only when the auction is open. We ran our segmented
invariant-confluence decision procedure on this example, and it was able to
deduce without any human interaction that the example was segmented
invariant-confluent in less than a tenth of a second.

\example[Escrow Transactions]\examplelabel{EscrowTransactionsEval}
Escrow transactions are a concurrency control technique that allows a database
to execute transactions that increment and decrement numeric values with more
concurrency than is otherwise possible with general-purpose techniques like
two-phase locking~\cite{o1986escrow}. The main idea is that a portion of the
numeric value is put in escrow, after which a transaction can freely decrement
the value so long as it is not decremented by more than the amount that has
been escrowed. We now show how segmented invariant-confluence can be used to
implement escrow transactions.

Consider again the PN-Counter $s = (p_1, p_2, p_3), (n_1, n_2, n_3)$ from
\exampleref{CounreachableExample} replicated on three servers with transactions
to increment and decrement the PN-Counter. In
\exampleref{CounreachableExample}, we found that concurrent decrements violate
invariant-confluence which led us to a segmentation which prohibited concurrent
decrements. We now propose a new segmentation with escrow amount $k$ that will
allow us to perform concurrent decrements that respect the escrowed value. The
first segment $(\setst{(p_1, p_2, p_3), (n_1, n_2, n_3)}{p_1, p_2, p_3 \geq k
\land n_1, n_2, n_3 \leq k}, T)$ allows for concurrent increments and
decrements so long as every $p_i \geq k$ and every $n_i \leq k$. Intuitively,
this segment represents the situation in which every server has escrowed a
value of $k$. They can decrement freely, so long as they don't exceed their
escrow budget of $k$. The second segment is the one presented in
\exampleref{CounreachableExample} which prohibits concurrent decrements. We ran
our decision procedure on this example and it concluded that it was segmented
invariant-confluent in less than a tenth of a second and without any human
interaction.

\textbf{TPC-C and Feral Concurrency Control.}

TODO

\subsection{Segmented Invariant Confluence}%
\seclabel{SegmentedInvariantConfluenceEval}
Now, we evaluate the performance of replicating an object with segmented
invariant-confluence as compared to replicating it with eventual consistency or
linearizability. We begin with two benchmarks that demonstrate the same
concept: the performance of segmented invariant-confluent replication varies
with the amount of global coordination induced by either (a) performing a
transaction that is disallowed within a segment or (b) transitioning between
segments.
